---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# openmpp

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![R-CMD-check](https://github.com/oncology-outcomes/openmpp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/oncology-outcomes/openmpp/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `openmpp` is to provide a programmatic interface to the OpenM++ API directly from R to simplify creating scenarios, running models, and gathering results for further processing.

## Installation

You can install the development version of `openmpp` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("oncology-outcomes/openmpp")
```

## Usage

The `openmpp` package contains many functions that provide access to nearly every OpenM++ API endpoint. However, users will typically only use a smaller set of functions for most common tasks.

### User Authentication

Each user is required to set their local or remote host address (i.e., URL) for the OpenM++ API in their global or project-specific `.Renviron` file in order for the `openmpp` package to authenticate and communicate with the API on behalf of the user. 

For an API running locally, set the following environment variable in your `.Renviron` file:

```
OPENMPP_LOCAL_URL=http://localhost:XXXX
```

Where `XXXX` is the four digits corresponding to your specific local host address. If you aren't sure of your host address, you may contact the OpenM++ administrator to retrieve this information.

For an API running remotely, set the following environment variables in your `.Renviron` file:

```
OPENMPP_REMOTE_URL=...
OPENMPP_REMOTE_USER=...
OPENMPP_REMOTE_PWD=...
```

Note that the URL, user name, and password should be kept confidential and not committed into version control (e.g., git).

Once the environment variables are set, users may register a local or remote API connection in their R scripts.

```r
library(openmpp)
use_OpenMpp_local()
```

Or,

```r
library(openmpp)
use_OpenMpp_remote()
```

see `?use_OpenMpp_local` or `?use_OpenMpp_remote` for more information.

### Main Functions

- Functions for accessing tables of models, worksets, or model runs
    
    + `get_models()`
    
    + `get_worksets()` / `get_scenarios()`
    
    + `get_model_runs()` / `get_runs()`

- Functions for creating new worksets or scenarios

    + `create_scenario()` / `create_workset()`
    
- Functions for loading models, worksets, or model runs

    + `load_model()`
    
    + `load_workset()` / `load_scenario()`
    
    + `load_model_run()` / `load_run()`
    
    + `load_model_runs()` / `load_runs()`

- Functions for deleting worksets or model runs

    + `delete_workset()` / `delete_scenario()`
    
    + `delete_model_run()` / `delete_run()`

### Models, Scenarios, Runs, and RunSets

There are 4 main classes you will work with when using the `openmpp` package: `OpenMppModel`, `OpenMppWorkset`, `OpenMppModelRun`, and `OpenMppModelRunSet`. Each of these are `R6` classes. `R6` is an encapsulated object-oriented programming system for R. Use the `load_*()` set of functions to load a model, workset/scenario, model run, or set of model runs into memory.

Instances of each of these 4 classes have methods and fields associated with them. You can access public methods and fields using the standard `$` subset operator (e.g., `obj$action()` or `obj$field`)

### Example

Here we will work through a very simple example of creating a new scenario, extracting parameters to change, changing parameters, running the model, and extracting results.

```{r message=FALSE}
library(openmpp)
library(dplyr)
library(ggplot2)

use_OpenMpp_remote()
```

Let's see what models are available:

```{r}
get_models()
```

We can now see what worksets and model runs exist for a given model.

```{r}
get_worksets('RiskPaths')
```

```{r}
get_runs('RiskPaths')
```

Now we can load the `RiskPaths` model to inspect.

```{r}
rp <- load_model('RiskPaths')
rp
```

We will now load the `Default` set of input parameters for the RiskPaths model.

```{r}
rp_default <- load_scenario('RiskPaths', 'Default')
rp_default
```

Finally, we will load the base run for the RiskPaths model.

```{r warning=FALSE}
baserun_digest <- rp$ModelRuns$RunDigest[[1]]
rp_baserun <- load_run('RiskPaths', baserun_digest)
rp_baserun
```

We will create a new scenario based on the parameters from the `RiskPaths_Default` model run.

```{r}
create_scenario('RiskPaths', 'MyNewScenario', baserun_digest)
```

We will load the new scenario, copy over the `AgeBaselinePreg1` parameter from the base run.

```{r warning=FALSE}
my_scenario <- load_scenario('RiskPaths', 'MyNewScenario')
```

Let's reduce the fertility rate by 10% across all age groups...

```{r}
my_scenario$copy_params('AgeBaselinePreg1')
```

```{r}
half_rate <- my_scenario$Parameters$AgeBaselinePreg1

half_rate <-
  half_rate |> 
  mutate(across(-sub_id, \(x) x * 0.9))

my_scenario$Parameters$AgeBaselinePreg1 <- half_rate
```

We will now run the model and give it the name `'ExampleRun'`. We use the `wait = TRUE` flag to make sure we want for the model run to finish before returning to our R session. We use `progress = FALSE` to avoid printing progress bars in this document. Note that model runs may take a long time when the number of simulation cases is large.

```{r}
my_scenario$ReadOnly <- TRUE
my_scenario$run('ExampleRun', wait = TRUE, progress = FALSE)
```

Note that we can use the `opts` argument and the `opts_run()` function to configure our run. By default, models are run with 5,000 simulation cases and 12 SubValues. This allows for quick model runs and faster iteration, but users will want to increase the number of simulation cases when performing a full model run.

Now that our model run is complete, let's load it into memory.

```{r warning=FALSE}
example_run <- load_run('RiskPaths', 'ExampleRun')
example_run
```

We can now extract an output table from the `Tables` field in the model run object (`example_run$Tables`).

```{r}
example_run$Tables$T06_BirthsByUnion
```

Great, we have created a new scenario, modified some parameters, ran the model, and extracted output tables. In this last step, we will load multiple model runs into memory to compare them.

```{r warning=FALSE}
rp_runs <- load_runs('RiskPaths', rp$ModelRuns$RunDigest)
rp_runs
```

We will extract a new table from both models. Note that an extra column, `RunName` is added to indicate which model run the output table data corresponds to.

```{r}
births <- rp_runs$Tables$T06_BirthsByUnion
births
```

We can even plot this using `ggplot2`! Note that the number of simulation cases for `ExampleRun` is **low** so the results are not to be trusted! This is only for demonstration purposes.

```{r warning=FALSE}
births |> 
  ggplot(aes(Dim0, expr_value, fill = RunName)) +
  geom_col(position = position_dodge()) +
  labs(x = NULL, y = 'Number of births by union') +
  coord_flip() +
  theme_minimal() +
  theme(legend.position = 'bottom')
```

When we are sure we no longer need a scenario or model run, we can use `delete_scenario()` or `delete_run()` to clean things up!

```{r echo=FALSE}
my_scenario$ReadOnly <- FALSE
delete_scenario('RiskPaths', 'MyNewScenario')
delete_model_run('RiskPaths', 'ExampleRun')
```

## Code of Conduct

Please note that the `openmpp` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
